---
title: "MS 204 Lab 3 solutions"
author: "PUT YOUR NAME HERE"
output: html_document
---


### Exercise 1

Use the appropriate function (dbinom, pbinom, rbinom) to find the probability that a binomial random variable `X` is less than 7 given that the number of trials is 20, and p=.4.

```{r}
pbinom(6, 20, 0.4)
```

**The probability that `X` is less than 7 given 20 trials with p = 0.4 is about 0.25.**

### Exercise 2

Which is more likely to happen? 

- A. You flip a coin 10 times and get exactly 5 heads

- B. You flip a coin 100 times and get exactly 50 heads

- C. You flip a coin 1000 times and get exactly 500 heads

Use either `dbinom`, `rbinom` or `pbinom` (whichever is appropriate here) to find out.  Are you suprised by the answer?

```{r}
dbinom(5, 10, 0.5)
dbinom(50, 100, 0.5)
dbinom(500, 1000, 0.5)
```

**The probabilities are 0.25 (10 tosses, 5 heads), 0.08 (100 tosses, 500 heads), and 0.025 (1000 tosses, 500 heads). I'm not surprised, but then again I'm your teacher so I shouldn't be surprised. Are you?**

### Exercise 3


**The barplot is symmetric when p is roughly 0.5. Skewness increases as p becomes further away from 0.5 (towards 0 or 1)**

### Exercise 4

This exercise is very similar to what we did for Henry and his burned pancakes.

Here we are learning about a Surgeon named Dr. Frankenfurter.  Dr Frankenfurter performs 9 surgeries a day.  Dr. Frankenfurter is an excellent surgeon, but even excellent surgeons sometimes make mistakes.  Let's say that Dr. Frankenfurter has a .008 chance of making a mistake for each surgery, and that surgeries are independent of each other.  

#### Exercise 4a

Simulate the random variable X, where X = the number of times that Dr. F makes a mistake in a surgery in a day, keeping in mind he performs 9 surgeries each day.

```{r}
mylist<-c(rep(1, 8), rep(0,992))
Bern.Trials.9<-sample(mylist, 9, replace=TRUE)
X<-sum(Bern.Trials.9)
X
```




#### Exercise 4b

Simulate 10000 days of Dr. Frankenfurter performing 9 surgeries a day.  Make a barplot of the number of mistakes he makes each day (as we did with Henry and his burned pancakes):

```{r}
set.seed(0)
number.mistakes<-NULL #making an empty list
n.days<-10000  #number of days Henry makes pancakes

for (i in 1:n.days){

  Bern.Trials.9<-sample(mylist, 9, replace=TRUE)
  X<-sum(Bern.Trials.9)  #Adding up the 4 Independent Bernoulli Trials to get my Binomial variable
  number.mistakes<-c(number.mistakes, X)
}

table(number.mistakes)/n.days

barplot(table(number.mistakes),names.arg=c(0,1,2), main="Number of Pancakes Henry Burned", col="chocolate4")  
```

In our simulation, he made 0 mistakes on 92.9 percent of days, 1 mistake on 6.78 percent of days, and 2 mistakes on 0.25 percent of days.

#### Exercise 4c

Based on your results in 4b, how likely is Dr. Frankfurter to make 2 or more mistakes in a day? Verify your results using either `dbinom`, `rbinom` or `pbinom.`

**Our simulations found 2 or more mistakes on 0.12 percent of days. We can get a similar fraction using `pbinom()`**

```{r}
1 - pbinom(1, 9, 0.008)
```

**The true probability of 2 or more mistakes is 0.0022, which roughly resembles our simulated findings (0.0025)**

### Exercise 5

On an assembly line, there is a 1% chance that a manufactured component is defective. Every day, 100 parts come off the assembly line. Find the probability that at least two parts are defective using the following

- (i)	Simulating using the R command rbinom(x,n,p)
- (ii) The R command dbinom and/or pbinom
- (iii)	Without using technology, by writing an expression for the desired result

```{r}
rand.draw <- rbinom(1000, 100, 0.01)
mean(rand.draw >= 2)
```

**(i)** By simulation, about 26.5 percent (answers may vary).

```{r}
1 - pbinom(1, 100, 0.01)
```

**(ii)** By using `pbinom()` as above, about 26.4 percent.

**(iii)** As entered into R, `1 - (100)*(0.01)^1*(0.99)^99 - (0.01)^0*(0.99)^100` = 0.264