---
title: "MA 204 Ch. 6 Lab"
author: "YOUR NAME"
--- 

Open a new R Markdown file (File / New File / R Markdown...).  Set the output to HTML mode, and click "Knit HTML" at the top (or Command+Shift+K on Mac).  This should produce a web page with the knitting procedure executing your code blocks.  You can edit this new file to produce final versions of your lab.

**It may be helpful to use the Lab06.Rmd file on Github as a template for your submission**.  Feel free to do so.  You access the raw code by clicking on the `raw` button.

Notice how the text you write in the .Rmd file shows up in the output file each time you click "Knit HTML".  

***General instructions for labs***: 

+  Open the R Markdown file (.Rmd) from the course Github page
+  Change the heading to include your author name
+  Save the R Markdown file (named as:  [MikeID]-[Lab01].Rmd -- e.g. "mlopez-Lab01.Rmd") to somewhere where you'll be able to access it later (zip drive, My Documents, Dropbox, etc)
+  Your file should contain the code/commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file
+  Each answer must be supported by written statements (unless otherwise specified) as well as any code used

### The apply function

One cool thing about r is that it includes many time saving functions.  Let's say I had a matrix a 30000x5 matrix and I wanted to find out the median of each row in the matrix filled with random normally distributed variables:

```{r}
rnormalz<-rnorm((30000*5))  #generating the numbers
rn.matrix<-matrix(rnormalz, ncol=5)  #putting the numbers in a matrix with 5 collumns
```

To find the median in each row I could write a loop like this:

```{r}
start <- proc.time()
medians <- NULL #making a vector to put my 300000 medians in
for(i in 1: 30000){
    medians <- c(medians, median(rn.matrix[i,]))  #takes the median of the ith row
}
end <- proc.time()

end[3] - start[3]
```

This loop takes a little while to run!  Let's use something called the apply function to make this go faster!

with the the apply function, I pass in three arguments:
1.  the matrix I am interested in
2.  whether I want to look at rows (pass in 1) or collumns (pass in 2)
3.  the function I want to perform on each row or collumn

Let's try the apply function:

```{r}
start <- proc.time()
medians <- apply(rn.matrix, 1, median) # takes the median of each row of the rn.matrix 
end <- proc.time()
end[3] - start[3]
```

That was much faster!! We will get a chance to use the apply function today.

Here are some other examples of ways to use the apply function:

```{r}

means.row<-apply(rn.matrix, 1, mean) # takes the mean of each row of the rn.matrix matrix

means.col<-apply(rn.matrix, 2, mean) #takes the mean of each collumn of the rn.matrix matrix

sums.col<-apply(rn.matrix, 2, sum) #takes the sum of each collumn of the rn.matrix matrix

```



##### Back to our regularly scheduled probability

Today we will get to look at Expectations and Variances, and how we can use R to double check our answers.

### Properties of expectations and variances

E(aX+b) = aE(X)+b, 
E(X+Y) = E(X)+E(Y), 

V(X) = E[ (X-E(X))^2 ], 
V(X) = E(X^2 ) -[E(X)^2 ], 
V(aX+b) = aV(X)

### Variance

Let's Simulate an exponential random variable with lambda parameter=.2 many many times:

```{r}
X <- rexp(100000, .2)

plot(density(X), col="blue", lwd=2, main="X~exp(.2)")
```


1. Check that the mean of `X` matches what you expected it to be, and confirm that this appears appropriate in the plot. 

The variance of an exponential(0.2) = 1/(0.2)^2

Let's use our many many exponentials to see if we get something close to this

```{r}
1/(0.2)^2
mean(X^2) - (mean(X))^2 #variance
var(X) #also the variance
```

2. Let Y = -2X + 4, which you can store in R by using `Y <- -2*X + 4`. Find the mean and variance of `Y` by hand, and use R to confirm your answers. 

```{r}

```

#### Independence


Now let's let X ~ exp(.2), Y ~ exp(.4), and X and Y are Independent.

We want to simulate Var(3x-4Y+2):

```{r}
X<-rexp(100000, .2)
Y<-rexp(100000, .4)

function.X.Y<-(3*X - 4*Y + 2)

var(function.X.Y) #Var(3X-4Y+2)

plot(density(function.X.Y), col="blue", lwd=2, main="3x-4Y+2")
```



Let's compare our result to  Var(3X) - Var(4y) + Var(2)

```{r}
part.1 = var(3*X)         #Var(3X)

part.2 = var(4*Y)         #Var(4Y)

part.3 <- mean((2)^2) - (mean(2))^2 #Var(2)

part.1 - part.2 + part.3

```

3. What's wrong with the above? Make it right


4. Simulate the E(X+Y) when X ~ exponential(3), Y ~ exponential(3), and confirm that your answer matches results when solving by hand 

```{r}

```


5.  Simulate the Variance(X+Y) when X~exponential(3), Y~exponential(3)

```{r}

```

6.  Plot the simulated pdf of (X+Y), when X~exponential(3), Y~exponential(3)

```{r}

```

