---
title: "MA 204 Sixth Lab"
author: "YOU AND YOUR PARTNER'S NAMES"
output: word_document
---


### The apply function

One cool thing about r is that it includes many time saving functions.  Let's say I had a matrix a 30000x5 matrix and I wanted to find out the median of each row in the matrix filled with random normally distributed variables:

```{r}
rnormalz<-rnorm((30000*5))  #generating the numbers
rn.matrix<-matrix(rnormalz, ncol=5)  #putting the numbers in a matrix with 5 collumns
```

To find the median in each row I could write a loop like this:

```{r}
start<-proc.time()
medians<-NULL #making a vector to put my 300000 medians in
for(i in 1: 30000){
    medians<-c(medians, median(rn.matrix[i,]))  #takes the median of the ith row
}
end<-proc.time()

end[3]-start[3]
```

This loop takes a little while to run!  Let's use something called the apply function to make this go faster!

with the the apply function, I pass in three arguments:
1.  the matrix I am interested in
2.  whether I want to look at rows (pass in 1) or collumns (pass in 2)
3.  the function I want to perform on each row or collumn

Let's try the apply function:
```{r}
start<-proc.time()
medians<-apply(rn.matrix, 1, median) # takes the median of each row of the rn.matrix 
end<-proc.time()
end[3]-start[3]
```

That was much faster!! We will get a chance to use the apply function today.

Here are some other examples of ways to use the apply function:
```{r}

means.row<-apply(rn.matrix, 1, mean) # takes the mean of each row of the rn.matrix matrix

means.col<-apply(rn.matrix, 2, mean) #takes the mean of each collumn of the rn.matrix matrix

sums.col<-apply(rn.matrix, 2, sum) #takes the sum of each collumn of the rn.matrix matrix

```



##### Back to our regularly scheduled probability

Today we will get to look at Expectations and Variances and how we can use R to double check our answers!!

### Properties of expectations and variances

E(aX+b)= aE(X)+b, 
E(X+Y)=E(X)+E(Y), 

V(X)= E[ (X-E(X))^2 ], 
V(X)=E(X^2 ) -[E(X)^2 ], 
V(aX+b)=aV(X)

### Variance

Let's Simulate an exponential random variable with lambda parameter=.2 many many times:

```{r}
X<-rexp(100000, .2)

plot(density(X), col="blue", lwd=2, main="X~exp(.2)")
```


The variance of an exponentail(.2)= 1/(.2)^2
Let's use our many many exponentials to see if we get something close to this

```{r}
1/(.2)^2
mean(X^2)-(mean(X))^2 #variance
var(X) #also the variance
```

Now let's let X~exp(.2), Y~exp(.4), and X and Y are Independent.

We want to simulate Var(3x-4Y+2):

```{r}
X<-rexp(100000, .2)
Y<-rexp(100000, .4)

function.X.Y<-(3*X -4*Y+2)

var(function.X.Y) #Var(3X-4Y+2)

plot(density(function.X.Y), col="blue", lwd=2, main="3x-4Y+2")
```



Let's compare our result to  Var(3X)-Var(4y)+Var(2)
```{r}
part.1=var(3*X)         #Var(3X)

part.2=var(4*Y)         #Var(4Y)

part.3<-mean((2)^2)-(mean(2))^2 #Var(2)

part.1-part.2+part.3

```

What's wrong here? Make it right


#### Exercise 1a
Simulate the E(X+Y) when X~exponential(3), Y~exponential(3)

```{r}

```


#### Exercise 1b
Simulate the Variance(X+Y) when X~exponential(3), Y~exponential(3)

```{r}

```

#### Exercise 1c
Plot the simulated pdf of (X+Y), when X~exponential(3), Y~exponential(3)

```{r}

```




### Geometric probabilities

A point (X,Y) is uniformly distributed on the triangle with vertices (0,0), (1,0), and (1,1). We are going to simulate the covariance and correlation between X and Y. 

```{r}

xsim<-c()
ysim<-c()
for (i in 1:10000){
  x<-runif(1)
  y<-runif(1)
  if (y<x){
    xsim<-c(xsim,x)
    ysim<-c(ysim,y)}}
cov(xsim,ysim)
cor(xsim,ysim)
```

#### Exercise 2a

Draw this triangle on a separate piece of paper. What is the area of this triangle?

####  Exercise 2b

If a point is randomly chosen on space S, then the joint density function for X and Y is simply 1/A(S), where A(S) is the area of S. What is the joint pdf in the triangle example?  

####  Exercise 2c

What are the constraints for x and y in this joint pdf? Use the Rcode and your picture to guide you.


####  Exercise 2d 

Find E[XY] using your joint pdf above. Double check in R using `mean(xsim*ysim)`

####  Exercise 2e

Find the marginal density of X and Y, as well as E[X] and E[Y]. Double check each of these using R (you can use `hist(xsim)` and `mean(xsim)`, for example) 



